---
title: "Assignment - Linear Mixed Effect"
author: "Dr Mohd Azmi P-UD 0079/19"
date: '`r format(Sys.Date(), "%d %B, %Y")`'
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float:
      collapsed: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
options(qwraps2_markup = "markdown", knitr.kable.NA = '')
getOption("qwraps2_frmt_digits", 2)
```

# Motivation

Diabetes control important to delay the complication of diabetes. In Malaysia, HbA1c have been used as one of parameter to monitor the diabetes control. In Malaysia, National Diabetes Registry (NDR) have been develop to monitor diabetes control among Malaysians population.

In typical linear regression, researcher would like to see the effect (or relationship) of predictors and outcome. However, in some situation, the sample had several levels in hierarchical forms. For example, researcher might want to see the relationship between venous fasting blood sugar (fbs) and HbA1c level on diabetic patient. Typically, the researcher can sample several diabetic patients and measure their fbs and HbA1c. However, these patient might have group effect, where they might have difference variance depending on their group. Patients in location A might have different value of relationship between fbs and hba1c, comparing with patients in location B. Multilevel model allow researcher to measure these multilevel variance. 

For this report, the data taken from NDR Perak, which consist of these data

1. district of the patient
2. patient current age
3. Ethnicity
4. Duration of diagnosed with Diabetes Mellitus
5. Patient's body weight in kilogram
6. Fasting Blood Sugar level (fbs)
7. HbA1c level

The research would like to see the relationship between fbs and HbA1c level. the researcher expected that the variance will be different for each district, thus multilevel model was used to estimate the relationship.


# Analysis

## Packages

These are the R's packages that used in this report.

```{r}
library(pacman) # package manager, elegant way to load packages
p_load(knitr, tidyverse, haven, summarytools, qwraps2, scales, lme4, broom, merTools)
```

## Dataset & Data Exploration

Due to very small estimate, data transformation need to be done for variable weight and the duration of dm. Weight were measure in per 10kg unit while the duration is in per 5 years of duration.

below is the dataset.

```{r}
lmeds <-  read_sav("mlm_perakndr.sav") %>%
  mutate_if(is.labelled, ~(as_factor(.))) %>%
  na.omit() %>%
  mutate(wt10 = Weight/10, dmdur5 = DMDur/5)
lmeds
```

Data exploration done to ensure no missing data nor data error.

```{r, eval = F}
descr(lmeds)
freq(lmeds)
```

The demographic and distribution of the data were summarised using following code

```{r}
demo_sum <- list("Daerah" = list("Bagan Datuk" = ~ n_perc0(.data$Daerah_ID == "Bagan Datuk"),
                                 "Batak Rabit" = ~ n_perc0(.data$Daerah_ID == "Batak Rabit"),
                                 "Changkat Jong" = ~ n_perc0(.data$Daerah_ID == "Changkat Jong"),
                                 "Chikus" = ~ n_perc0(.data$Daerah_ID == "Chikus"),
                                 "Degong" = ~ n_perc0(.data$Daerah_ID == "Degong"),
                                 "Durian Sebatang" = ~ n_perc0(.data$Daerah_ID == "Durian Sebatang"),
                                 "Selama" = ~ n_perc0(.data$Daerah_ID == "Selama"),
                                 "Jenderata" = ~ n_perc0(.data$Daerah_ID == "Jenderata"),
                                 "Langkap" = ~ n_perc0(.data$Daerah_ID == "Langkap"),
                                 "Selekoh" = ~ n_perc0(.data$Daerah_ID == "Selekoh"),
                                 "Simpang Empat" = ~ n_perc0(.data$Daerah_ID == "Simpang Empat"),
                                 "Rungkup" = ~ n_perc0(.data$Daerah_ID == "Rungkup"),
                                 "Larut" = ~ n_perc0(.data$Daerah_ID == "Larut"),
                                 "Teluk Intan" = ~ n_perc0(.data$Daerah_ID == "Teluk Intan"),
                                 "Kampar" = ~ n_perc0(.data$Daerah_ID == "Kampar"),
                                 "Gopeng" = ~ n_perc0(.data$Daerah_ID == "Gopeng"),
                                 "Malim Nawar" = ~ n_perc0(.data$Daerah_ID == "Malim Nawar"),
                                 "Kerian" = ~ n_perc0(.data$Daerah_ID == "Kerian"),
                                 "Kinta" = ~ n_perc0(.data$Daerah_ID == "Kinta")),
                 "Age (year)" = list("Mean (SD)" = ~ mean_sd(.data$Age, denote_sd = "paren")),
                 "Gender" = list("Female" = ~ n_perc0(.data$GenCode == "Female"),
                                 "Male" = ~ n_perc0(.data$GenCode == "Male")),
                 "Ethnicity" = list("Malay" = ~ n_perc0(.data$EthCode == "Malay"),
                                    "Chinese" = ~ n_perc0(.data$EthCode == "Chinese"),
                                    "Indian" = ~ n_perc0(.data$EthCode == "Indian"),
                                    "Other" = ~ n_perc0(.data$EthCode == "Other")),
                 "Duration of DM (years)" = list("Mean (SD)" = ~ mean_sd(.data$DMDur, denote_sd = "paren")),
                 "Height" = list("Mean (SD)" = ~ mean_sd(.data$Height, denote_sd = "paren")),
                 "Weight" = list("Mean (SD)" = ~ mean_sd(.data$Weight, denote_sd = "paren")),
                 "Fasting Blood Sugar (mmol/L)" = list("Mean (SD)" = ~ mean_sd(.data$fbsa, denote_sd = "paren")),
                 "HbA1c (%)" = list("Mean (SD)" = ~ mean_sd(.data$hbaa, denote_sd = "paren")))

meanhbabydis<- lmeds %>% group_by(Daerah_ID) %>% summarise(n = n(), `HbA1c Mean` = mean(hbaa))
```





# Analysis

## Linear Regression

Analysis was started with typical linear regression, to estimate the regression parameter between ethnicity, fasting blood sugar level, weight (in 10kg unit) and duration of DM (in 5 years unit), regardless of the variances at district level.

```{r}
lrmod <- lm(hbaa ~ EthCode + fbsa + wt10 + dmdur5, lmeds)
```


## Linear Mixed Effect

Since the researcher expected the variance between the districts will be varies, linear mixed model should be done to measure/adjust for the variance.

### Null Model

Null model was measured as baseline for comparison with linear mixed models with random intercept and random slopes. In null model, the mean of HbA1c will be estimated, regardless of the covariates, with consideration of district level variance.

```{r}
lmemodnull <- lmer(hbaa ~ 1 + (1 | Daerah_ID), data = lmeds, REML = F)
```

The variance between district and within can be measured.

```{r}
lmemodnull_var <- VarCorr(lmemodnull)
```

### Random Intercept Model

In Random Intercept model, the HbA1c was estimated by considering the effect of covariates (in this model include ethnicity, fasting blood sugar level, weight in 10kg unit and duration of DM in 5 year unit), and district level variance.

```{r}
lmemodRI <- lmer(hbaa ~ EthCode + fbsa + wt10 + dmdur5 + (1 | Daerah_ID), data = lmeds, REML = F)
```

The variance between district and within can be measured.

```{r}
lmemodRI_var <- VarCorr(lmemodRI)
```


### Random Slope Model

Random slope model was done to check whether there are also effect of district variance in the covariate. In this study, the researcher wanted to know does fasting blood sugar level had different effect by different district.

```{r}
lmemodRS <- lmer(hbaa ~ EthCode + fbsa + wt10 + dmdur5 + (1 + fbsa | Daerah_ID), lmeds, control = lmerControl(optimizer = "bobyqa"), REML = F)
```

The variance between district and within can be measured.

```{r}
lmemodRS_var <- VarCorr(lmemodRS)
```

Comparison between Random Slope Model and Random Intercept was done with likelihood ratio test (LR Test)

```{r}
comp_RIRS <- anova(lmemodRI, lmemodRS)
```

Scatter plot will be use to visualize the effect of random slope

```{r}
lmemodRS_ranef <- ranef(lmemodRS, condVar = T)$Daerah_ID %>% mutate(intercept = `(Intercept)`, slope = fbsa)
lmemodRS_ranef_plot <- ggplot(lmemodRS_ranef, aes(x = intercept, y = slope)) + geom_point() + geom_vline(xintercept = 0) + geom_hline(yintercept = 0)

```

Random slope can also be visualize with forest plot.

```{r}
lmemodRS_resim <- REsim(lmemodRS)
lmemodRS_interceptplot <- lmemodRS_resim %>%
  filter(term == "(Intercept)") %>%
  ggplot(aes(x = reorder(groupID, median), y = median, ymin = median - 1.96*sd, ymax = median + 1.96*sd)) +
  geom_pointrange() + coord_flip() + ylab("(Intercept)") + xlab("Daerah")
lmemodRS_slopeplot <-lmemodRS_resim %>%
  filter(term == "fbsa") %>%
  ggplot(aes(x = reorder(groupID, median), y = median, ymin = median - 1.96*sd, ymax = median + 1.96*sd)) +
  geom_pointrange() + coord_flip()+ ylab("Slope") + xlab("Daerah")
```


### Model Fitness

Scatterplot was constructed between expected and residual HbA1C. This was to visualized the linearity and homoscedascity of the residual.

```{r}
resid_plot <- augment(lmemodRS) %>%
  ggplot(aes(x=.fitted, y=.resid, colour = Daerah_ID)) +
  geom_point() +
  geom_hline(yintercept = c(-5,0,5), linetype = 2) +
  xlab("Expected HbA1c") + ylab("Residual")
```

# Result

## Demographic

Below are the summary of the data

```{r, results='asis'}
print(summary_table(lmeds, demo_sum), rtitle = "Variables", cnames = c("Mean (SD) / n (%)"))
```

Mean HbA1c for each district are as below

```{r}
meanhbabydis
```

## Linear Regression

```{r}
summary(lrmod)
```

## Linear Mixed Model

### Null Model

```{r}
summary(lmemodnull)
```

The variance between district and within district for null model were as below

```{r}
print(lmemodnull_var, comp = "Variance", digits =2)
```

Between Daerah variance: `r round(as.data.frame(VarCorr(lmemodnull))[1,4],2)`
Within School variance: `r round(as.data.frame(VarCorr(lmemodnull))[2,4],2)`

In null model, proportion of unexplained variance due to differences between Daerah is `r round(as.data.frame(VarCorr(lmemodnull))[1,4],2)` / (`r round(as.data.frame(VarCorr(lmemodnull))[1,4],2)` + `r round(as.data.frame(VarCorr(lmemodnull))[2,4],2)`) = `r round((as.data.frame(VarCorr(lmemodnull))[1,4] / (as.data.frame(VarCorr(lmemodnull))[1,4] + as.data.frame(VarCorr(lmemodnull))[2,4]))*100,2)`%


### Random Intercept Model

```{r}
summary(lmemodRI)
```

The variance between district and within district were as below

```{r}
print(lmemodRI_var, comp = "Variance", digits =2)
```

Between Daerah variance: `r round(as.data.frame(VarCorr(lmemodRI))[1,4],2)`
Within School variance: `r round(as.data.frame(VarCorr(lmemodRI))[2,4],2)`

In RI model, proportion of unexplained variance due to differences between Daerah is `r round(as.data.frame(VarCorr(lmemodRI))[1,4],2)` / (`r round(as.data.frame(VarCorr(lmemodRI))[1,4],2)` + `r round(as.data.frame(VarCorr(lmemodRI))[2,4],2)`) = `r round((as.data.frame(VarCorr(lmemodRI))[1,4] / (as.data.frame(VarCorr(lmemodRI))[1,4] + as.data.frame(VarCorr(lmemodRI))[2,4]))*100,2)`%



### Random Slope Model

```{r}
summary(lmemodRS)
```

The variance between district and within district were as below

```{r}
print(lmemodRS_var, comp = "Variance", digits =2)
```

Between Daerah variance: `r round(as.data.frame(VarCorr(lmemodRS))[1,4],2)`
Within School variance: `r round(as.data.frame(VarCorr(lmemodRS))[4,4],2)`

In RS model, proportion of unexplained variance due to differences between Daerah is`r round(as.data.frame(VarCorr(lmemodRS))[1,4],2)` / (`r round(as.data.frame(VarCorr(lmemodRS))[1,4],2)` + `r round(as.data.frame(VarCorr(lmemodRS))[4,4],2)`) =  `r round((as.data.frame(VarCorr(lmemodRS))[1,4] / (as.data.frame(VarCorr(lmemodRS))[1,4] + as.data.frame(VarCorr(lmemodRS))[4,4]))*100,2)`%


```{r}
comp_RIRS
```

Comparison with LR test show that the random slope model was significant, and thus will be used as final model in this study.

Random effect can be visualize as below.

```{r}
lmemodRS_ranef_plot
```

The scatter plot show that there are some district that have lower intercept (lower mean HbA1c), but higher effect of the slope (fasting blood sugar) - at the left upper of the plot, and conversely there are some district that have higher intercept (higer mean HbA1c), but lower effect of the slope (fasting blood sugar) - at the right lower of the plot.

```{r}
lmemodRS_interceptplot
lmemodRS_slopeplot
```


### Model Assessment

Scatter plot between expected and residual was shown below

```{r}
resid_plot
```
The scatter plot show that the plot scattered uniformly scattered at residual = 0, indicating that the assumption of linearity and homoscedascity were met.

# Discussion

# Conclusion

# Reference


# Additional Info

Session Info

```{r}
sessionInfo()
```