---
title: "Assignment - Machine Learning (Classification))"
author: "Dr Mohd Azmi P-UD 0079/19"
date: '`r format(Sys.Date(), "%d %B, %Y")`'
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
options(qwraps2_markup = "markdown", knitr.kable.NA = '')
```

# Motivation

Predict which notification will get delayed.

# Environment

## Packages

```{r}
library(pacman) # package manager, elegant way to load packages
p_load(tidyverse, haven, DT, rpart, summarytools, broom, caret, neuralnet)
set.seed(119)
```

## Dataset & Data Exploration

```{r}
delnotids <- read_sav("classdsfin2.sav") %>% mutate_if(is.labelled, ~(as_factor(.)))
#delnotids <- delnotids0[sample(47295, 5000),]
#delnotids %>% select(-id) %>% datatable()
```

```{r}

sRow <- sample(nrow(delnotids), nrow(delnotids)*.7)

delnotids_train <- delnotids[sRow,]
delnotids_test <- delnotids[-sRow,]
```


# Analysis & Result

## Descriptive

```{r}
freq(delnotids)
```




## Machine Learning - Decision Tree

### Decision Tree Model

Using train dataset

```{r}
delnoti_dtmod <- rpart(DelayCode ~ Disease + NotiReqCode + FaciCollapseCode + DxWeekdayRecode + CaseClassReCode + DxStatRecodeNoMissing + ACDPCDReCode + CaseRegisterCode + Region + QuarterCode,
                       data = delnotids_train,
                       method = "class",
                       parms = list(split = "information"),
                       control = rpart.control(cp=0.001))
plot(delnoti_dtmod)
text(delnoti_dtmod)
```

### Prediction

```{r}
delnotids_test_pred <- predict(delnoti_dtmod, newdata = delnotids_test, type = "class")
delnotids_test_predcomb <- tibble(delnotids_test, Predicted = delnotids_test_pred)
delnotids_test_predcomb %>% select(DelayCode, Predicted, Disease, NotiReqCode, FaciCollapseCode, DxWeekdayRecode, CaseClassReCode, DxStatRecodeNoMissing, ACDPCDReCode, CaseRegisterCode, Region, QuarterCode)
```



### Model Evaluation

using test dataset

```{r}
dteval_res <- confusionMatrix(delnotids_test_predcomb$Predicted, delnotids_test_predcomb$DelayCode, positive = "Yes")
dteval_res
```

The model have a mixture picture.

While the accuracy (or overall percentage) is good with `r round((dteval_res$overall['Accuracy'])*100,1)`% correctly classify the outcome and good specificity (able to predict outcome "no") of `r round((dteval_res$byClass['Specificity'])*100,1)`%, but the sensitivity was low at `r round((dteval_res$byClass['Sensitivity'])*100,1)`% only.

## Machine Learning - Artificial Neural Network

```{r}
delnotidsNum <- read_sav("classdsfin2.sav")
delnotidsNum_train <- delnotidsNum[sRow,]
delnotidsNum_test <- delnotidsNum[-sRow,]


delnoti_annmod <- neuralnet(DelayCode ~ Disease + NotiReqCode + FaciCollapseCode + DxWeekdayRecode + CaseClassReCode + DxStatRecodeNoMissing + ACDPCDReCode + CaseRegisterCode + Region + QuarterCode, data = delnotidsNum_train, hidden=1, linear.output = FALSE)
plot(delnoti_annmod)
```


# Discussion


# Conclusion

# Reference


# Additional Info

```{r}

```




```{r}
sessionInfo()
```

