---
title: "Assignment - Ordinal Logistic Regression"
author: "Dr Mohd Azmi P-UD 0079/19"
date: '`r format(Sys.Date(), "%d %B, %Y")`'
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
options(qwraps2_markup = "markdown", knitr.kable.NA = '')
```

# Motivation

Delayed notification will delay public health action.

In this example, the researcher 


# Environment

## Packages

```{r}
library(pacman) # package manager, elegant way to load packages
p_load(tidyverse, haven, DT, broom, summarytools, knitr, scales, ordinal, caret, VGAM, LogisticDx)
```

## Dataset & Data Exploration



```{r}
ordds <-  read_sav("classdsfin2.sav") %>% mutate_if(is.labelled, ~(as_factor(.)))
ordds
```


# Analysis & Result

## Descriptive


```{r, results="asis"}
freq(ordds)

```



```{r, results="asis"}
descr(ordds)
```

```{r}
ggplot(ordds, aes(Disease, fill = Disease)) +
  geom_bar() + scale_y_continuous(labels = comma)
ggplot(ordds, aes(NotiDurCode, fill = NotiDurCode)) +
  geom_bar() + scale_y_continuous(labels = comma)
ggplot(ordds, aes(FacilityCodeGC, fill = FacilityCodeGC)) +
  geom_bar() + scale_y_continuous(labels = comma) + coord_flip()
ggplot(ordds, aes(DxDayReCodeMon, fill = DxDayReCodeMon)) +
  geom_bar() + scale_y_continuous(labels = comma) + coord_flip()
ggplot(subset(ordds, !is.na(CaseClassReCode)), aes(CaseClassReCode, fill = CaseClassReCode)) +
  geom_bar() + scale_y_continuous(labels = comma)
ggplot(ordds, aes(DxStatRecodeNoMissing, fill = DxStatRecodeNoMissing)) +
  geom_bar() + scale_y_continuous(labels = comma)
ggplot(subset(ordds, !is.na(ACDPCDReCode)), aes(ACDPCDReCode, fill = ACDPCDReCode)) +
  geom_bar() + scale_y_continuous(labels = comma)
ggplot(ordds, aes(StateReCode, fill = StateReCode)) +
  geom_bar() + scale_y_continuous(labels = comma) + coord_flip()
ggplot(ordds, aes(DxMonthCode, fill = DxMonthCode)) +
  geom_bar() + scale_y_continuous(labels = comma) + coord_flip()
```


## Multinomial Logistic Regression

### Univariable

```{r}
levels(ordds$NotiDurCode)
mulnomds <- ordds %>% mutate(NotiDurCode2 = fct_relevel(NotiDurCode, levels = c("Late", "Expected", "Early")))
levels(mulnomds$NotiDurCode2)

mulnomlr_dis <- vglm(NotiDurCode2 ~ Disease, data=mulnomds, family = "multinomial")
summary(mulnomlr_dis)
```

### Multivariable

```{r}
mulnommlr <- vglm(NotiDurCode ~ FaciCollapseCode + DxWeekdayRecode + NotiReqCode + Region, data=mulnomds, family = "multinomial")
summary(mulnommlr)
```

### Model Checking

Logit 1 - Early vs Late

```{r}
levels(ordds$NotiDurCode)
mulnomlog1ds <- ordds %>% filter(NotiDurCode == c("Early", "Late"))

logreglog1 <- glm(NotiDurCode ~ FaciCollapseCode + DxWeekdayRecode + NotiReqCode + Region, data=mulnomlog1ds, family = "binomial")
logreglog1aug <- augment(logreglog1, type.predict = "response", type.residual = "response") %>%
  mutate(PredCat = cut(.fitted, breaks = c(0, 0.49999,1),
                       labels = c("Early", "Late")))
descr(logreglog1aug$.fitted)
confusionMatrix(logreglog1aug$PredCat, logreglog1aug$NotiDurCode, positive = "Late")
gof_logreglog1 <- gof(logreglog1, plotROC = T, g=12)
```

for model 1, accuracy was `r round(confusionMatrix(logreglog1aug$PredCat, logreglog1aug$NotiDurCode, positive = "Late")$overall[1]*100, 1)`% (95% CI = `r round(confusionMatrix(logreglog1aug$PredCat, logreglog1aug$NotiDurCode, positive = "Late")$overall[3]*100, 1)`%, `r round(confusionMatrix(logreglog1aug$PredCat, logreglog1aug$NotiDurCode, positive = "Late")$overall[4]*100, 1)`%) with `r round(confusionMatrix(logreglog1aug$PredCat, logreglog1aug$NotiDurCode, positive = "Late")$byClass[2]*100, 1)`% specificity but `r round(confusionMatrix(logreglog1aug$PredCat, logreglog1aug$NotiDurCode, positive = "Late")$byClass[1]*100, 1)`% sensitivity  

## Ordinal Logistic Regression

### Univariable

```{r}
ordlr_dis <- clm(NotiDurCode ~ FaciCollapseCode + DxWeekdayRecode + NotiReqCode + Region, data=ordds)
summary(ordlr_dis)
tidy(ordlr_dis, exponentiate = T, conf.int = T)
```



2. The odd of Measles being in expected or above is `r round(exp(coef(ordlr_dis)[3]), 2)` higher  `r round(exp(confint(ordlr_dis)[1,]), 2)`  (95% CI: 1.10, 1.46, p = 0.001) than a normal BMI.

### Multivariable

```{r}
ordmlr <- clm(NotiDurCode ~ FaciCollapseCode + DxWeekdayRecode + NotiReqCode + Region, data=ordds)
summary(ordmlr)
tidy(ordmlr, exponentiate = T, conf.int = T)
```


### Model Fitness

```{r}
ordmlr_predds <- augment(ordmlr, type.predict = "class")
```

```{r}
confusionMatrix(ordmlr_predds$.fitted, ordmlr_predds$NotiDurCode)
```


### Assumption

```{r}
nominal_test(ordmlr)
```

###


# Discussion



# Conclusion

# Reference


# Additional Info

Session Info

```{r}
sessionInfo()
```

