---
title: "Assignment - Ordinal Logistic Regression"
author: "Dr Mohd Azmi P-UD 0079/19"
date: '`r format(Sys.Date(), "%d %B, %Y")`'
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
options(qwraps2_markup = "markdown", knitr.kable.NA = '')
```

# Motivation

Delayed notification will delay public health action.

In this example, the researcher 


# Environment

## Packages

```{r}
library(pacman) # package manager, elegant way to load packages
p_load(tidyverse, haven, DT, broom, summarytools, surveytoolbox, knitr, blandr)
```

## Dataset & Data Exploration

In this data, there are 5 variables

1. id : ID for each district
2. ppka : PPKA status - presence of PPKA in the district (0 - no PPKA, 1 - PPKA present)
3. popnum : Number of population in the district - per 10,000 population
4. total : total disease notification received
5. delaynum : number of delayed notification received

```{r}
set.seed(100)

v_districtid <- c(1:100) %>% set_varl("District ID")
v_total <- rpois(100,8) %>% set_varl("No of Total Notification")

v_delay <- v_total %>%    #random where higher total will have higher delay 
  map_dbl(function(x){
  if(x >=11){
    sample(5:9, size = 1, prob = seq(10,30,5))
  } else if(x >= 6){
    sample(2:4, size = 1, prob = c(40,35,25))
  } else{
    sample(0:1, size = 1, prob = c(60,40))
  }
  }) %>% set_varl("No of Delayed Notification")

v_popnum <- v_total %>%     #random where higher total will have higher population
  map_dbl(function(x){
  if(x >=11){
    sample(60:80, size = 1)
  } else if(x >= 6){
    sample(40:60, size = 1)
  } else{
    sample(20:40, size = 1)
  }
  }) %>% set_varl("No of Population (hundred)")

v_ppka <- v_popnum %>%     #random yes no for ppka - bigger population have higher prob to have ppka
  map_dbl(function(x){
    if(x >= 65){
      sample(0:1, size = 1, prob = c(30,70))
    } else if(x >= 45){
      sample(0:1, size = 1, prob = c(40,60))
    } else{
      sample(0:1, size = 1, prob = c(50,50))
    }
  }) %>% 
  set_vall(value_labels = c("No PPKA" = 0, "PPKA Present" = 1)) %>% 
  set_varl("Present of PPKA")

deldisds <- tibble(id = v_districtid, ppka = v_ppka, pop = v_popnum/10, total = v_total, delaynum = v_delay-v_ppka*2) %>% filter(delaynum >= 0) %>% 
  mutate_if(is.labelled, ~(as_factor(.)))
datatable(deldisds,
          class = "cell-border stripe",
          colnames = c("DistrictID", "PPKA Status", "Population No. ('0,000)", "Total Notification", "No. of Delayed Notification"))
```


# Analysis & Result

## Descriptive

From 94 district, 60% of them have PPKA

```{r, results="asis"}
freq(deldisds$ppka)
ggplot(deldisds, aes(x=ppka, fill = ppka)) + geom_bar() + theme_bw() + xlab("PPKA Status")
```

The median population in each district is `r format(median(deldisds$pop)*10000, format="d", big.mark=",")` population. Below is the distribution of district for particular number of population

```{r, results="asis"}
descr(deldisds$pop, stats = c("mean", "sd", "med", "q1", "q3", "iqr"))
ggplot(deldisds, aes(x=pop)) + geom_bar() + scale_x_binned() + theme_bw() + xlab("No of Population ('0,000)") + ylab("No of District")
```

The median total disease notification in each district is `r median(deldisds$total)` notifications. Below is the distribution of district for total number of disease notification


```{r, results="asis"}
descr(deldisds$total, stats = c("mean", "sd", "med", "q1", "q3", "iqr"))
ggplot(deldisds, aes(x=total)) + geom_bar() + scale_x_binned() + theme_bw() + xlab("No of Disease Notification") + ylab("No of District")
```


The median delayed disease notification in each district is `r median(deldisds$delaynum)` notifications. Below is the distribution of district for total number of disease notification


```{r, results="asis"}
descr(deldisds$delaynum, stats = c("mean", "sd", "med", "q1", "q3", "iqr"))
ggplot(deldisds, aes(x=delaynum)) + geom_bar() + scale_x_binned() + theme_bw() + xlab("No of Delayed Disease Notification") + ylab("No of District")
```

Scatterplot below show both total notification and delayed notification increase when the population is increase.

```{r}
ggplot(deldisds, aes(x=pop)) +
  geom_point(aes(y=total, colour = "red")) + geom_smooth(aes(y=total, colour = "red")) +
  geom_point(aes(y=delaynum, colour = "blue")) + geom_smooth(aes(y=delaynum, colour = "blue")) +
  scale_x_binned() + theme_bw() + xlab("No of Population ('0,000)") + ylab("No of Total Notification \n& Delayed Notification")
```


## Poisson Linear Model

### Model Estimation


Poisson Regression shows that both population variable and present of PPKA were signficant

```{r}
delay_mod <- glm(delaynum ~ pop + ppka + offset(log(total)), family = poisson, data = deldisds)
summary(delay_mod)
tidy(delay_mod, conf.int = T)  %>%
  mutate(coef = estimate, IRR = exp(estimate), IRRllci = exp(conf.low), IRRulci = exp(conf.high)) %>% select(term, coef, std.error, IRR, IRRllci, IRRulci, p.value) %>% kable()
```

For each increase of 1 unit of population (i.e. 10,000 population), the log rate of delayed notification per total notification was increase by `r round(coef(delay_mod)[2],2)`, which equivalence to IRR of `r round(exp(coef(delay_mod)[2]),2)` (95% CI = `r round(exp(confint(delay_mod)[2,]),2)`).

Among district with PPKA, the log rate of delayed notification per total notification was `r round(coef(delay_mod)[3],2)` (i.e. those with PPKA had lower rate) as compare to district without PPKA. The IRR of rate delayed notification per total notification for district with PPKA as compared to those without PPKA was `r round(exp(coef(delay_mod)[3]),2)` (95% CI = `r round(exp(confint(delay_mod)[3,]),2)`).

```{r}
tidy(delay_mod, conf.int=T) %>% 
  ggplot(aes(x=exp(estimate), y=term, color = term)) +
    geom_point() +
    geom_errorbarh(aes(xmin = exp(conf.low), xmax = exp(conf.high))) +
    geom_vline(lty=2, xintercept = 1) + xlab("IRR")
```


## Observed vs Predicted

Below are the predicted number of delayed notification as compared to observed delayed notification for each of sample.

```{r}
augm_delaymod <- augment(delay_mod, type.predict = "response", type.resid = "response") %>% 
  mutate(total = exp(offset.log.total..)) %>% 
  select(ppka, pop, total, delaynum, .fitted, .resid)
datatable(augm_delaymod, colnames = c("PPKA Status", "Population No. ('0,000)", "Total Notification", "Observed No. of Delayed Notification", "Predicted No. of Delayed Notification", "Residual"))
```

Agreement between observed and predicted were measure using Bland Altman Plot

```{r}
p_load(blandr)

BAdata_ph <- blandr.statistics(augm_delaymod$delaynum, augm_delaymod$.fitted, sig.level = 0.95)
blandr.plot.ggplot(BAdata_ph, method1name = "Observed", method2name = "Predicted", ciDisplay = T) +
  scale_x_continuous(name = "Average Observed & Predicted", breaks = seq(0,10,1)) +
  scale_y_continuous(name = "Different Observed & Predicted") +
  annotate(geom="text", label="+ 1.96 SD = 2.49", x=6.5, y = 2.7) +
  annotate(geom="text", label="- 1.96 SD = - 2.49", x=6.5, y = -2.2) +
  annotate(geom="text", label="Mean Diff = 0.00", x= 6.5, y = 0.3) + theme_bw() +
  ggtitle("Bland Altman Plot for Observed vs Predicted")
```

Bland Altman Plot show the difference between observed and predicted count of delayed notification were on average 0 difference with 95% agreement between +/- 2.5 of the different count.  

### Prediction

Both expected rate and expected count for delayed notification was calculated using the model, with these parameter

1. PPKA present vs No PPKA
2. 30,000 vs 70,000 population
3. 5 vs 10 total notification

```{r}
newds2 <- expand.grid(ppka = c("No PPKA", "PPKA Present"), pop = c(3,7), total = c(5,10))
augment(delay_mod, type.predict = "response", newdata = newds2) %>%
  select(-.se.fit) %>%
  mutate(rate = .fitted/total) %>% 
  datatable(class = "cell-border stripe",
            colnames = c("PPKA Status", "Population No. ('0,000)", "Total Notification", "Expected/Predicted No. of Delayed Notification", "Expected Rate (Expected No / Total)"))

```

# Discussion

- Present of PPKA reduce the rate / count of delayed notification
- Why poisson rate

# Conclusion

# Reference


# Additional Info

Session Info

```{r}
sessionInfo()
```

