---
title: "Assignment - Survival"
author: "Mohd Azmi"
date: "06/07/2020"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
options(qwraps2_markup = "markdown", knitr.kable.NA = '')
```

# Motivation

Delayed notification will delay public health action.

In this example, the researcher would like to measure / predict number of delayed notification.

Note: Data was generated (Refer [Dataset](#dataset-data-exploration))

# Environment

## Packages

```{r}
library(pacman) # package manager, elegant way to load packages
p_load(tidyverse, haven, DT)
```

## Dataset & Data Exploration

```{r}
p_load(surveytoolbox)
set.seed(100)

v_id <- seq(1, 10000) %>% set_varl("Record Identifier")

v_state <- sample(0:4, size = 10000, replace = T, prob = seq(0.3, 0.7, 0.1)) %>%
  set_vall(value_labels = c("Kel" = 0, "Sab" = 1, "Ked" = 2, "Sel" = 3, "Joh" = 4)) %>%
  set_varl("State")

v_disease <- v_state %>%
  map_dbl(function(x){
    if(x == 3){
      sample(0:4, size = 1, prob = c(10, 12, 20, 23, 35))
    } else if(x == 1){
      sample(0:4, size = 1, prob = c(10, 27, 20, 23, 20))
    } else {
      sample(0:4, size = 1, prob = c(10, 17, 20, 23, 30)) 
    }
  }) %>%
  set_vall(value_labels = c("MERS" = 0, "Mala" = 1, "Meas" = 2, "HFMD" = 3, "Deng" = 4)) %>%
  set_varl("Disease")

v_delay <- v_disease %>%
  map_dbl(function(x){
    if(x == 4){
      sample(0:1, size = 1, prob = c(80,20))
    } else if(x == 3){
      sample(0:1, size = 1, prob = c(60, 40))
    } else {
      sample(0:1, size = 1, prob = c(70, 30)) 
    }
  }) %>%
  set_vall(value_labels = c("No" = 0, "Yes" = 1)) %>%
  set_varl("DelayStat")

deldisds <- tibble(id = v_id, state=v_state, disease = v_disease, delaystat = v_delay)
deldisds %>% mutate_if(is.labelled, ~(as_factor(.))) %>% select(-id) %>% datatable()
```


# Analysis & Result

## Descriptive

### Sociodemographic


## Survival Analysis


### Prediction


# Discussion


# Conclusion

# Reference


# Additional Info

```{r}

```




```{r}
sessionInfo()
```

