---
title: "Assignment - Survival"
author: "Mohd Azmi"
date: "06/07/2020"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
options(qwraps2_markup = "markdown", knitr.kable.NA = '')
```

# Motivation

Delayed notification will delay public health action.

In this example, the researcher would like to measure / predict number of delayed notification.

Note: Data was generated (Refer [Dataset](#dataset-data-exploration))

# Environment

## Packages

```{r}
library(pacman) # package manager, elegant way to load packages
p_load(tidyverse, haven, DT)
```

## Dataset & Data Exploration

```{r}
p_load(surveytoolbox)
set.seed(100)

v_districtid <- c(1:100) %>% set_varl("District ID")
v_total <- rpois(100,8) %>% set_varl("No of Total Notification")

v_delay <- v_total %>% 
  map_dbl(function(x){
  if(x >=11){
    sample(5:9, size = 1, prob = seq(10,30,5))
  } else if(x >= 6){
    sample(2:4, size = 1, prob = c(40,35,25))
  } else{
    sample(0:1, size = 1, prob = c(60,40))
  }
  }) %>% set_varl("No of Delayed Notification")

v_popnum <- v_total %>% 
  map_dbl(function(x){
  if(x >=11){
    sample(60:80, size = 1)
  } else if(x >= 6){
    sample(40:60, size = 1)
  } else{
    sample(20:40, size = 1)
  }
  }) %>% set_varl("No of Population (hundred)")

v_ppka <- v_popnum %>% 
  map_dbl(function(x){
    if(x >= 65){
      sample(0:1, size = 1, prob = c(70,30))
    } else if(x >= 45){
      sample(0:1, size = 1, prob = c(50,50))
    } else{
      sample(0:1, size = 1, prob = c(30,70))
    }
  }) %>% 
  set_vall(value_labels = c("No PPKA" = 1, "PPKA Present" = 0)) %>% 
  set_varl("Present of PPKA")

deldisds <- tibble(id = v_districtid, ppka = v_ppka, pop = v_popnum, total = v_total, delaynum = v_delay) %>% mutate_if(is.labelled, ~(as_factor(.)))
datatable(deldisds,
          class = "cell-border stripe",
          colnames = c("DistrictID", "PPKA Status", "Population No. ('000)", "Total Notification", "No. of Delayed Notification"))
```


# Analysis & Result

## Descriptive

### Sociodemographic


## Survival Analysis


### Prediction


# Discussion


# Conclusion

# Reference


# Additional Info

```{r}

```




```{r}
sessionInfo()
```

