---
title: "Assignment - Structural Equation Modelling"
author: "Dr Mohd Azmi P-UD 0079/19"
date: '`r format(Sys.Date(), "%d %B, %Y")`'
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
options(qwraps2_markup = "markdown", knitr.kable.NA = '')
```


# Motivation

https://doi.org/10.1371/journal.pone.0230644


# Environment

## Packages

```{r}
library(pacman) # package manager, elegant way to load packages
p_load(haven, tidyverse)  
       
```

## Dataset & Data Exploration

Data was saved in SPSS save file, imported to R

```{r}
dbmds <- read_sav("pone.0230644.s002.sav") %>% mutate_if(is.labelled, ~(as_factor(.)))
dbmds
```


# Analysis & Result

## Descriptive

### Sociodemographic

Descriptive analysis was done on sociodemographic data, was shown below. 

```{r, results='asis'}
#demo_sum <- list("Gender" = list("Male" = ~ n_perc0(.data$Gender == "Male"),
#                                 "Female" = ~ n_perc0(.data$Gender == "Female")),
#                 "Age (years)" = list("Mean (SD)" = ~ mean_sd(.data$Age, denote_sd = "paren")),
#                 "Ethnicity" = list("Malay" = ~ n_perc0(.data$Race == "Malay"),
#                                    "Non-Malay" = ~ n_perc0(.data$Race == "Non-Malay")),
#                 "comorbidities (status)" = list("Yes" = ~ n_perc0(.data$CM == "Yes"),
#                                               "No" = ~ n_perc0(.data$CM == "No")),
#                 "DM (status)" = list("Yes" = ~ n_perc0(.data$CMDM == "DM"),
#                                      "No" = ~ n_perc0(.data$CMDM == "No")),
#                 "Hypertension (status)" = list("Yes" = ~ n_perc0(.data$CMHPT == "HPT"),
#                                                "No" = ~ n_perc0(.data$CMHPT == "No")),
#                 "Ischaemic Heart Disease (status)" = list("Yes" = ~ n_perc0(.data$CMIHD == "IHD"),
#                                                           "No" = ~ n_perc0(.data$CMIHD == "No")))
#print(summary_table(pdcds1, demo_sum), rtitle = "Variables", cnames = c("Mean (SD) / n (%)"))
```


### Score - Outcomes

out of all the samples, more than half of them underwent percutaneous method, with half of them develop complication (mean complication time = 107 days) and eventually 46% of all catheter require removal (mean removal time = 117 days).

```{r, results='asis'}

```


## Structural Equation Model


# Discussion



# Conclusion

# Reference


# Additional Info

```{r}

```




```{r}
sessionInfo()
```

