---
title: "Assignment - Structural Equation Modelling"
author: "Dr Mohd Azmi P-UD 0079/19"
date: '`r format(Sys.Date(), "%d %B, %Y")`'
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
options(qwraps2_markup = "markdown", knitr.kable.NA = '')
```


# Motivation

Every student will encounter difficulty during their study. In addition to being student, MMED candidate also need to do clinical works, taking care of their patient, which is part of their study. This will put extra strain to the student.

There are several questionnaire tools that can be use to measure both 1) the level of stress, and 2) specific cause of stress.

PSQ have been use to measure what are the domain that causing stress (stressor domains), which include

1. academic stressor
2. work-family
3. job prospect

One of common tools that use to measure stress score is PSS. while the final score is actually only one domain, but for this assignment purpose, the score was group into positive score (questions that will have high score for high stress), and negative score (questionas tha will have low score for high stress).

In this assignment, the research would like to measure the effect of different stressors to the stress score.

SEM allow researcher to find the complex relationship between exogenous and endogenous variables, both observed and latent variables.


# Environment

## Packages

```{r}
library(pacman) # package manager, elegant way to load packages
p_load(haven, tidyverse, semTools, semPlot, DT)  
       
```

## Dataset & Data Exploration

Data was saved in SPSS save file, imported to R

```{r}
stressds <- read_sav("SEMstress.sav") %>% mutate_if(is.labelled, ~(as_factor(.)))
stressds %>% datatable(fillContainer = T)
```


# Analysis & Result

## Descriptive

### Sociodemographic

Descriptive analysis was done on sociodemographic data, was shown below. 

```{r, results='asis'}
#demo_sum <- list("Gender" = list("Male" = ~ n_perc0(.data$Gender == "Male"),
#                                 "Female" = ~ n_perc0(.data$Gender == "Female")),
#                 "Age (years)" = list("Mean (SD)" = ~ mean_sd(.data$Age, denote_sd = "paren")),
#                 "Ethnicity" = list("Malay" = ~ n_perc0(.data$Race == "Malay"),
#                                    "Non-Malay" = ~ n_perc0(.data$Race == "Non-Malay")),
#                 "comorbidities (status)" = list("Yes" = ~ n_perc0(.data$CM == "Yes"),
#                                               "No" = ~ n_perc0(.data$CM == "No")),
#                 "DM (status)" = list("Yes" = ~ n_perc0(.data$CMDM == "DM"),
#                                      "No" = ~ n_perc0(.data$CMDM == "No")),
#                 "Hypertension (status)" = list("Yes" = ~ n_perc0(.data$CMHPT == "HPT"),
#                                                "No" = ~ n_perc0(.data$CMHPT == "No")),
#                 "Ischaemic Heart Disease (status)" = list("Yes" = ~ n_perc0(.data$CMIHD == "IHD"),
#                                                           "No" = ~ n_perc0(.data$CMIHD == "No")))
#print(summary_table(pdcds1, demo_sum), rtitle = "Variables", cnames = c("Mean (SD) / n (%)"))
```


### Score - Outcomes



```{r, results='asis'}

```


## Structural Equation Model


In this dataset, the variables are

1. 3 latent independent factors
  + Academic Domain, which measured by 4 independent observed variables
    - PSQ_Q1:
    - PSQ_Q11
    - PSQ_Q20
    - PSQ_Q28
  + Work-family domain
    - PSQ_Q2s
    - PSQ_Q8s
    - PSQ_Q16s
    - PSQ_Q24s
  + Performance pressure
    - PSQ_Q7s
    - PSQ_Q13s
    - PSQ_Q15s
    - PSQ_Q22s
2. 2 latent dependent factors
  + Positive Score, which measured by 6 dependent observed variables
    - PSS_Q1s
    - PSS_Q2s
    - PSS_Q3s
    - PSS_Q6s
    - PSS_Q9s
    - PSS_Q10s
  + Negative Score, which measured by 4 dependent observed variables
    - PSS_Q4s
    - PSS_Q5s
    - PSS_Q7s
    - PSS_Q8s


```{r}
modful <- "
posstr =~ PSS_Q1s + PSS_Q2s + PSS_Q3s + PSS_Q6s + PSS_Q9s + PSS_Q10s
negstr =~ PSS_Q4s + PSS_Q5s + PSS_Q7s + PSS_Q8s
acad =~ PSQ_Q1s + PSQ_Q11s + PSQ_Q20s + PSQ_Q28s
wfam =~ PSQ_Q2s + PSQ_Q8s + PSQ_Q16s + PSQ_Q24s
perf =~ PSQ_Q7s + PSQ_Q13s + PSQ_Q15s + PSQ_Q22s
posstr + negstr ~ acad + wfam + perf

"

sem_modful <- sem(modful, data = stressds)
```


the plot (with standardized estimate shown) is as below

```{r}
semPaths(sem_modful, what = "par", whatLabels = "std", edge.color = "black", layout = "tree2", residuals = F, rotation = 2)
```

in this model, the unstandardized regression coefficient for academic on positive score was `r round(sem_modful_res$PE [23,5],2)`, which mean for each increase of 1 unit of academic stressor, there will be increase of `r round(sem_modful_res$PE [23,5],2)` unit of positive stress score. the standardized regression coefficeint was `r round(sem_modful_res$PE [23,10],2)` - which mean for every increase of 1 standard deviation of academic stressor latent variable, there will be increase of `r round(sem_modful_res$PE[23,10],2)` standard deviation of positive stress latent score variable.

the full estimate (unstandardized coefficient), and r-square value as below

```{r}
summary(sem_modful, fit.measures = F, standardized = T, rsquare = T, header = F)
```

the standardized estimate as below

```{r}
standardizedSolution(sem_modful) %>% datatable() %>% formatRound(4:10, digits = 3) 
```


### Model Fit Assessment

```{r}
fitMeasures(sem_modful, c("chisq", "df", "pvalue", "cfi", "tli", "srmr", "rmsea"), output = "text")
```

CFI was low, however other model fit parameter show the model was fit.

```{r}
modindices(sem_modful) %>% filter(mi >= 3.86)
```

there are several path that have modification indexes higher than 3.84, however since this is confirmatory model, no modification will be done to the SEM path.



# Discussion



# Conclusion

# Reference


# Additional Info

```{r}

```




```{r}
sessionInfo()
```

