---
title: "Assignment - Survival"
author: "Mohd Azmi"
date: "06/07/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Environment

## Packages

```{r}
library(pacman) # package manager, elegant way to load packages
p_load(haven, #import spss data
       tidyverse,
       naniar, #replace values with missing
       lubridate #date manipulation
       )  
       
```

## Dataset

Data was saved in SPSS save file, imported to R and wrangling was done

1. select related variables
2. replace NA in variable for dateofremoval, for those with catheter not removed


```{r}
pdcds0 <- read_sav("PDC Survival - Saifullah 290819.sav")
pdcds0 %>% mutate_if(is.labelled, ~(as_factor(.))) %>% arrange(DataID) %>% print()
```


DataID, GenderCode, RaceCode, Age, MethodCode, Dateofplacement, Dateofremoval, diffday, Comorbidities, ComorbidYesNo, CMDM, CMHPT, CMIHD, CxCode, TypeofComplication, CxEarlyLate14Code, CxEarlyLate14TypeCode, CxEarlyLate30Code, CxEarly30YesNo, CxLate30YesNo, CxEarlyLate30TypeCode, ExitSiteIfxCode

```{r}
pdcds1 <- pdcds0 %>%
  mutate_if(is.labelled, ~(as_factor(.))) %>%
  arrange(DataID) %>% 
  mutate(Dateofremoval2 = as.character(Dateofremoval)) %>% 
  replace_with_na(replace = list(Dateofremoval2 = "2173-10-13")) %>%
  mutate(Dateofremoval = as.Date(Dateofremoval2)) %>%
  mutate(diffday = as.double(difftime(ymd(Dateofremoval), ymd(Dateofplacement), units = "days"))) %>% 
  select(DataID, GenderCode, RaceCode, Age, MethodCode, Dateofplacement, Dateofremoval, diffday, ComorbidYesNo, CMDM, CMHPT, CMIHD,CatheterStatusCode, RemovalReasonCode) 
pdcds1
```


# Analysis

## Data exploration
