---
title: "Assignment - Survival"
author: "Mohd Azmi"
date: "06/07/2020"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, results = c("asis"))
```

# Environment

## Packages

```{r}
library(pacman) # package manager, elegant way to load packages
p_load(haven, #import spss data
       tidyverse,
       naniar, #replace values with missing
       lubridate, #date manipulation
       summarytools, #data exploration
       DT, #create beautiful table
       survival,
       broom,
       knitr,
       qwraps2)  
       
```

## Dataset & Data Exploration

Data was saved in SPSS save file, imported to R

```{r}
pdcds0 <- read_sav("pdcsurvivalalter.sav")
pdcds0 %>% 
  arrange(DataID) %>%
  mutate_if(is.labelled, ~(as_factor(.))) %>%
  select(-c(DataID, Comorbidities)) %>%
  datatable(fillContainer = T)
```

data wrangling was done

1. select related variables
2. replace NA for those without recorded date for complication and removal of catheter
3. replace 365 days for duration with no complication nor removal of catheter (as the patient was follow up for maximum 1 year duration)



```{r}
pdcds1 <- pdcds0 %>%
  mutate_if(is.labelled, ~(as_factor(.))) %>%
  arrange(DataID) %>% 
  mutate(RemovalDate2 = as.character(RemovalDate),
         CxDate2 = as.character(CxDate)) %>% 
  replace_with_na(replace = list(RemovalDate2 = "2173-10-13",
                                 CxDate2 = "2173-10-13")) %>%
  mutate(RemovalDate = as.Date(RemovalDate2),
         CxDate = as.Date(CxDate2)) %>%
  mutate(RemovalDur = as.double(difftime(ymd(RemovalDate), ymd(PlaceDate), units = "days")),
         CxDur = as.double(difftime(ymd(CxDate), ymd(PlaceDate), units = "days"))) %>%
  replace_na(list(CxDur = 365, RemovalDur = 365)) %>%
  select(DataID, Gender, Race, Age, CM, CMDM, CMHPT, CMIHD, Method, PlaceDate, Compx, CxTypeCode, CxDate, CxDur, CatheterStatus, RemovalDate, RemovalDur, RemovalReason)

pdcds1 %>% select(-DataID) %>% datatable(fillContainer = T)
```

Simple data exploration to ensure necessary data wrangling was done.

```{r}
pdcds1 %>% select(Age) %>% summary() %>% kable()
pdcds1 %>% select_if(is.factor) %>% select(-c(CxTypeCode, RemovalReason)) %>% summary() %>% kable()
pdcds1 %>% select(CxTypeCode, RemovalReason) %>% summary() %>% kable()
```

# Analysis

## Descriptive

### Sociodemographic

```{r}
demo_sum <- list("Gender" = list("Male" = ~ n_perc0(.data$Gender == "Male"),
                                 "Female" = ~ n_perc0(.data$Gender == "Female")),
                 "Age (years)" = list("Mean (SD)" = ~ mean_sd(.data$Age, denote_sd = "paren")),
                 "Ethnicity" = list("Malay" = ~ n_perc0(.data$Race == "Malay"),
                                    "Non-Malay" = ~ n_perc0(.data$Race == "Non-Malay")))
print(summary_table(pdcds1, demo_sum), rtitle = "Demographic Summaries", cnames = c("Mean (SD) / n (%)"))
```


### Outcome

Among all the participants, there are 105 samples that developed catheter related, ranging from the earliest 13 days post catheterization, and longest reported after 339 days post catheterization.

```{r}
pdcds1 %>% filter(RemovalDur != 365) %>% descr(RemovalDur, stats = c("min", "max", "n.valid"))
```



```{r}
removaldurdt <- pdcds1 %>% group_by(RemovalDur) %>% count(RemovalDur)
ggplot(removaldurdt, aes(x=RemovalDur, y=n)) +
  geom_point() + geom_line() +
  scale_x_continuous(limits = c(0,364), name = "Duration of Complication (day)") +
  scale_y_continuous(limits = c(0,5), name = "No. of Patient")
```

No of catheter remove developed, by number of days

```{r}
cxdurdt <- pdcds1 %>% group_by(CxDur) %>% count(CxDur)
ggplot(cxdurdt, aes(x=CxDur, y=n)) +
  geom_point() + geom_line() +
  scale_x_continuous(limits = c(0,364), name = "Duration of Complication (day)") +
  scale_y_continuous(limits = c(0,5), name = "No. of Patient")
```

```{r}

```

## Survival Analysis

### Univariable

```{r}
cxsurvexpmod_gender <- survreg(Surv(CxDur, Compx == "Yes") ~ Gender, data = pdcds1, dist = "exponential")
summary(cxsurvexpmod_gender)
```

```{r}
cxsurvexpmod_method <- survreg(Surv(CxDur, Compx == "Yes") ~ Method, data = pdcds1, dist = "exponential")
summary(cxsurvexpmod_method)
```


Both time-to-event for complication and removal of catheter were analysed using survival analysis
- ethnicity and cm status not included



```{r}
cxsurvexpmod <- survreg(Surv(CxDur, Compx == "Yes") ~ Gender + Age + CMDM + Method, data = pdcds1, dist = "exponential")
summary(cxsurvexpmod)
```
a more elegant way to present it

```{r}
tidy(cxsurvexpmod) %>% 
  mutate(TR = exp(estimate),
         llci = exp(conf.low),
         ulci = exp(conf.high),
         HR = exp(-estimate)) %>% 
  select(term, estimate, TR, llci, ulci, p.value, HR)
```



# Additional Info

```{r}

```




```{r}
sessionInfo()
```

