---
title: "Assignment - Survival"
author: "Mohd Azmi"
date: "06/07/2020"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
options(qwraps2_markup = "markdown", knitr.kable.NA = '')
```

# Motivation

Peritoneal dialysis (PD) is one of renal replacement therapy available for end stage renal failure. However PD require tenckhoff catheterization to access to peritoneal fluid, which have their own risks. There are several technique for tenckhoff catheter placement. In HRPZ II, commonly tenckhoff catheterization was done with either percutaneous or peritonesocopic method.

There are several complications related to tenckhoff catheterization, which generally categorized into infection or mechanical complication. In this analysis, there are several objectives that the researcher would like to achieve

1. the time-to-complication and time-to-removal of the  tenckhoff catheter.
2. compare between percutaneous and peritonescopic method
3. find factors that may affect the time-to-complication and time-to-removal of the  tenckhoff catheter.

Note: 

1. while this research and dataset were based on ongoing research by one of MMed student, the dataset was altered.
2. hence any findings and interpretation from this analysis does not reflect the real outcome of patients in HRPZ and SHOULD NOT be use other than this assignment.

# Environment

## Packages

```{r}
library(pacman) # package manager, elegant way to load packages
p_load(haven, #import spss data
       tidyverse,
       naniar, #replace values with missing
       lubridate, #date manipulation
       summarytools, #data exploration
       DT, #create beautiful data table
       survival, #survival analysis
       broom, #tidy function
       knitr, #knit rmd
       qwraps2, #create beautiful descriptive statistic
       SurvRegCensCov,
       survminer)  
       
```

## Dataset & Data Exploration

Data was saved in SPSS save file, imported to R

```{r}
pdcds0 <- read_sav("pdcsurvivalalter.sav")
pdcds0 %>% 
  arrange(DataID) %>%
  mutate_if(is.labelled, ~(as_factor(.))) %>%
  select(-c(DataID, Comorbidities)) %>%
  datatable(fillContainer = T)
```

data wrangling was done

1. select related variables
2. replace NA for those without recorded date for complication and removal of catheter
3. replace 365 days for duration with no complication nor removal of catheter (as the patient was follow up for maximum 1 year duration)


```{r}
pdcds1 <- pdcds0 %>%
  mutate_if(is.labelled, ~(as_factor(.))) %>%
  arrange(DataID) %>% 
  mutate(RemovalDate2 = as.character(RemovalDate),
         CxDate2 = as.character(CxDate)) %>% 
  replace_with_na(replace = list(RemovalDate2 = "2173-10-13",
                                 CxDate2 = "2173-10-13")) %>%
  mutate(RemovalDate = as.Date(RemovalDate2),
         CxDate = as.Date(CxDate2)) %>%
  mutate(RemovalDur = as.double(difftime(ymd(RemovalDate), ymd(PlaceDate), units = "days")),
         CxDur = as.double(difftime(ymd(CxDate), ymd(PlaceDate), units = "days"))) %>%
  replace_na(list(CxDur = 365, RemovalDur = 365)) %>%
  select(DataID, Gender, Race, Age, CM, CMDM, CMHPT, CMIHD, Method, PlaceDate, Compx, CxTypeCode, CxDate, CxDur, CatheterStatus, RemovalDate, RemovalDur, RemovalReason)

pdcds1 %>% select(-DataID) %>% datatable(fillContainer = T)
```

Simple data exploration to ensure necessary data wrangling was done.

```{r}
pdcds1 %>% select(where(is.factor)) %>% summary() %>% kable()
pdcds1 %>% select(-where(is.factor)) %>% summary() %>% kable()
```

# Analysis & Result

## Descriptive

### Sociodemographic

Descriptive analysis was done on sociodemographic data, was shown below. 

```{r, results='asis'}
demo_sum <- list("Gender" = list("Male" = ~ n_perc0(.data$Gender == "Male"),
                                 "Female" = ~ n_perc0(.data$Gender == "Female")),
                 "Age (years)" = list("Mean (SD)" = ~ mean_sd(.data$Age, denote_sd = "paren")),
                 "Ethnicity" = list("Malay" = ~ n_perc0(.data$Race == "Malay"),
                                    "Non-Malay" = ~ n_perc0(.data$Race == "Non-Malay")),
                 "comorbidities (status)" = list("Yes" = ~ n_perc0(.data$CM == "Yes"),
                                               "No" = ~ n_perc0(.data$CM == "No")),
                 "DM (status)" = list("Yes" = ~ n_perc0(.data$CMDM == "DM"),
                                      "No" = ~ n_perc0(.data$CMDM == "No")),
                 "Hypertension (status)" = list("Yes" = ~ n_perc0(.data$CMHPT == "HPT"),
                                                "No" = ~ n_perc0(.data$CMHPT == "No")),
                 "Ischaemic Heart Disease (status)" = list("Yes" = ~ n_perc0(.data$CMIHD == "IHD"),
                                                           "No" = ~ n_perc0(.data$CMIHD == "No")))
print(summary_table(pdcds1, demo_sum), rtitle = "Variables", cnames = c("Mean (SD) / n (%)"))
```


### Catheterization and Outcomes

out of all the samples, more than half of them underwent percutaneous method, with half of them develop complication (mean complication time = 107 days) and eventually 46% of all catheter require removal (mean removal time = 117 days).

```{r, results='asis'}
pdcds1_withna <- pdcds1 %>%
  replace_with_na(replace = list(RemovalDur = "365", CxDur = "365",
                                 CxTypeCode = "No Complication", RemovalReason = "N/A - Not removed"))
method_sum <- list("Method" = list("Percutaneous Method" = ~ n_perc0(.data$Method == "Percutaneous Method", digits = 1),
                                 "Peritoneoscopy Method" = ~ n_perc0(.data$Method != "Percutaneous Method", digits = 1)),
                   "Complication" = list("Yes" = ~ n_perc0(.data$Compx == "Yes", digits = 1),
                                         "No" = ~ n_perc0(.data$Compx != "Yes", digits = 1)),
                   "Mean Complication Time (n=118)" = list("Mean (SD)" = ~ mean_sd(.data$CxDur, na_rm = T, show_n = "never",
                                                                           denote_sd = "paren", digits = 2)),
                   "Reason for Complication" = list("Infection Complication" = ~ n_perc0(.data$CxTypeCode == "Infection Complication",
                                                                                         na_rm = T, digits = 1),
                                                    "Mechanical Complication" = ~ n_perc0(.data$CxTypeCode == "Mechanical Complication",
                                                                                          na_rm = T, digits = 1),
                                                    "Both" = ~ n_perc0(.data$CxTypeCode == "Both",
                                                                       na_rm = T, digits = 1)),
                   "Removal" = list("Yes" = ~ n_perc0(.data$CatheterStatus != "censor", digits = 1),
                                    "No" = ~ n_perc0(.data$CatheterStatus == "censor", digits = 1)),
                   "Mean Removal Time (n=105)" = list("Mean (SD)" = ~ mean_sd(.data$RemovalDur, na_rm = T, show_n = "never",
                                                                      denote_sd = "paren", digits = 2)),
                   "Reason for Removal" = list("Peritonitis" = ~ n_perc0(.data$RemovalReason == "Peritonitis", na_rm = T, digits = 1),
                                               "Exit Site Infection" = ~ n_perc0(.data$RemovalReason == "Exit Site Infection",
                                                                                 na_rm = T, digits = 1),
                                               "Malfunction" = ~ n_perc0(.data$RemovalReason == "Malfunction", na_rm = T, digits = 1)))
print(summary_table(pdcds1_withna, method_sum), rtitle = "Variables", cnames = c("Mean (SD) / n (%)"))
```

Among all the participants, there are 118 samples that developed catheter related complication, ranging from the earliest 13 days post catheterization, and longest reported after 317 days post catheterization.

```{r, results='asis'}
pdcds1 %>% filter(CxDur != 365) %>% descr(CxDur, stats = c("min", "max"))
```

there are 1 to 7 samples developed complication per day. 

```{r}
cxdurdt <- pdcds1 %>% group_by(CxDur) %>% count(CxDur)

ggplot(cxdurdt, aes(x=CxDur, y=n)) +
  geom_point() + geom_line() +
  scale_x_continuous(limits = c(0,337.5), breaks = seq(0,350,25), expand = c(0,0), name = "Duration of Catheterization (day)") +
  scale_y_continuous(limits = c(0,8), expand = c(0,0), name = "No. of Patient develop Complication") +
  theme_bw() + ggtitle("Line plot for number of complication for each duration of catheterization")

ggplot(cxdurdt, aes(x=CxDur, y=cumsum(n))) +
  geom_point() + geom_line() +
  scale_x_continuous(limits = c(0,337.5), breaks = seq(0,350,25), expand = c(0,0), name = "Duration of Catheterization (day)") +
  scale_y_continuous(limits = c(0, 130), expand = c(0,0), name = "No. of Patient (Cumulative) develop Complication") +
  theme_bw() + ggtitle("Cummulative Frequency Plot for number of \ncomplication for each duration of catheterization")
```

Among all the participants, there were 105 catheter that require removal, rangin from as early as 13 days post catheterization until 339 days post catheterization.

```{r results='asis'}
pdcds1 %>% filter(RemovalDur != 365) %>% descr(RemovalDur, stats = c("min", "max"))
```

there are 1 to 5 catheter removed per day

```{r}
removaldurdt <- pdcds1 %>% group_by(RemovalDur) %>% count(RemovalDur)

ggplot(removaldurdt, aes(x=RemovalDur, y=n)) +
  geom_point() + geom_line() +
  scale_x_continuous(limits = c(0,364), breaks = seq(0,350,25), expand = c(0,0), name = "Duration of Catheterization (day)") +
  scale_y_continuous(limits = c(0,6), expand = c(0,0), name = "No. of Catheter Removed") +
  theme_bw() + ggtitle("Line plot for number of removal for each duration of catheterization")

ggplot(removaldurdt, aes(x=RemovalDur, y=cumsum(n))) +
  geom_point() + geom_line() +
  scale_x_continuous(limits = c(0,350), breaks = seq(0,350,25), expand = c(0,0), name = "Duration of Complication (day)") +
  scale_y_continuous(limits = c(0, 110), expand = c(0,0), name = "No. of Patient (Cumulative) \nrequire Catheter Removal") +
  theme_bw() + ggtitle("Cummulative Frequency Plot for number of \nCatheter Removed for each duration of catheterization")

```


## Survival Analysis

### Catheter Survival (Descriptive)

Time-to-event for complication and catheter removal were describe using kaplan-meier analysis.

Out of 132 patients on percutaneous method, 75 develop complication, with median time-to-event for complication was 220 days.
Meanwhile, out of 98 patients on peritoneoscopy method, 43 develop complication, with median time-to-event for complication was beyond 365 days.

```{r}
compx_km <- survfit(Surv(time = CxDur, Compx == "Yes") ~ Method,
                      data = pdcds1,
                      type = "kaplan-meier")
compx_km
```

Probability of survival (not develop complication) for percutaneous method for 6-months was 54.5% (95% CI = 46.7, 63.7) while peritonescopy was 62.2% (95% CI = 53.4, 72.6). The probability of survival (not develop complication) for 30-days, 3-months and 1-year was shown below.

```{r}
summary(compx_km, times = c(30, 90, 180, 365))
```

Out of 132 patients on percutaneous method, 70 require removal, with median time-to-event for removal was 276 days.
Meanwhile, out of 98 patients on peritoneoscopy method, 35 require removal, with median time-to-event for removal was beyond 365 days.

```{r}
removal_km <- survfit(Surv(time = RemovalDur, CatheterStatus == "Catheter Removed") ~ Method,
                      data = pdcds1,
                      type = "kaplan-meier")
removal_km
```

Probability of survival (not removed) for percutaneous method for 6-months was 59.8% (95% CI = 52.0, 68.8) while peritonescopy was 71.4% (95% CI = 63.0, 81.0). The probability of survival (not removed) for 30-days, 3-months and 1-year was shown below.

```{r}
summary(removal_km, times = c(30, 90, 180, 365))
```

KM Plot for catheter complication was shown below

```{r}
ggsurvplot(compx_km, data = pdcds1, pval = T) + ggtitle("KM Plot for catheter complication")
```

KM Plot for catheter removal was shown below

```{r}
ggsurvplot(removal_km, data = pdcds1, pval = T) + ggtitle("KM Plot for catheter removal")
```


## Parametric Survival Analysis - Catheter Complication

As the distribution of the outcome was unknown, Survival Analysis with Weibull model was done.

Due to small cells, variables ethnicity, presence of comorbidities, hypertension status and ischaemic heart disease status were excluded from analysis

### Univariable

Parametric Survival Analysis with Weibull Model was done to measure the time-to-event of catheter complication, stratified by method of catheterization, as shown below.

```{r}
cxsurvweimod_method <- survreg(Surv(CxDur, Compx == "Yes") ~ Method, data = pdcds1, dist = "weibull")
summary(cxsurvweimod_method)
ConvertWeibull(cxsurvweimod_method, conf.level = 0.95)

```

variable method was not significant, with time ratio (TR) of duration for catheter complication for peritonesocopy method was `r round(cxsurvweimod_method_res$ETR, 2)[,1]` (95% CI = `r round(cxsurvweimod_method_res$ETR, 2)[,2:3]`) times than the duration for catheter complication for percutaneous method. The hazard ratio (HR) of peritonesocopy method (as compare to percutaneous method) to develop catheter was `r round(cxsurvweimod_method_res$HR, 2)[,1]` (95% CI = `r round(cxsurvweimod_method_res$HR, 2)[,2:3]`) times higher.


```{r}
WeibullDiag(Surv(CxDur, Compx == "Yes") ~ Method, data = pdcds1)
```

Weibull model seems unfit for gender variable, as the lines were linear but not parallel.

```{r}
cxsurv_method_kmfit <- tidy(survfit(Surv(CxDur, Compx == "Yes") ~ Method, data = pdcds1))
ggplot(cxsurv_method_kmfit, aes(x= log(time), y = log(estimate/(1-estimate)), colour = strata)) +
  geom_point() + geom_line() +
  theme_bw() + ggtitle("PH Assumption Checking Plot") + theme(legend.position = c(0.8, 0.8))
```

The plot seems straight line and parallel, thus does not violated proportional hazard assumptions.


### Multivariable

Parametric Survival Analysis with Weibull model was done to find the difference between method, when adjusted to gender, age and diabetes mellitus status.

```{r}
cxsurvweimod <- survreg(Surv(CxDur, Compx == "Yes") ~ Gender + Age + CMDM + Method, data = pdcds1, dist = "weibull")
summary(cxsurvweimod)
ConvertWeibull(cxsurvweimod, conf.level = 0.95)
```

when adjusted to gender, age and diabetes status, there was no significant association between method and time-to-event of catheter complication, with time ratio (TR) of peritonesocopy method (as compare to percutaneous method) to develop catheter complication was `r round(cxsurvweimod_res$ETR, 2)[4,1]` (95% CI = `r round(cxsurvweimod_res$ETR, 2)[4,2:3]`). 

the hazard ratio (HR) of peritonesocopy method (as compare to percutaneous method) to develop catheter was `r round(cxsurvweimod_res$HR, 2)[4,1]` (95% CI = `r round(cxsurvweimod_res$HR, 2)[4,2:3]`) times higher, when adjusted to gender, age and diabetes status.

### Prediction

Prediction for Survival Time - Catheter Complication

```{r}
newds <- data.frame(expand.grid(Gender = "Male", Age = seq(40,60,10), CMDM = c("DM", "No"), Method = c("Percutaneous Method", "Peritoneoscopy Method")))
predmod <- predict(cxsurvweimod, newdata = newds, type = "quantile", p = c(.25,.5,.75))
data.frame(cbind(newds, predmod)) %>%
  mutate("Predicted Complication Time (Days) - Median" = X2,
         "Predicted Complication Time (Days) - Q1" = X1,
         "Predicted Complication Time (Days) - Q3" = X3) %>%
  select(Gender, Age, CMDM, Method, "Predicted Complication Time (Days) - Q1", "Predicted Complication Time (Days) - Median", "Predicted Complication Time (Days) - Q3")
```

Prediction for Survival Probability


## Parametric Survival Analysis - Catheter Removal

As the distribution of the outcome was unknown, Survival Analysis with Weibull model was done.

Due to small cells, variables ethnicity, presence of comorbidities, hypertension status and ischaemic heart disease status were excluded from analysis

### Univariable

Parametric Survival Analysis with Weibull Model was done to measure the time-to-event of catheter removal, stratified by method of catheterization, as shown below.

```{r}
removalsurvweimod_method <- survreg(Surv(RemovalDur, CatheterStatus == "Catheter Removed") ~ Method, data = pdcds1, dist = "weibull")
summary(removalsurvweimod_method)
ConvertWeibull(removalsurvweimod_method, conf.level = 0.95)
```

variable method was significant, with time ratio (TR) of duration for catheter removal for peritonesocopy method was `r round(removalsurvweimod_method_res$ETR, 2)[,1]` (95% CI = `r round(removalsurvweimod_method_res$ETR, 2)[,2:3]`) times than the duration for catheter removal for percutaneous method. The hazard ratio (HR) of peritonesocopy method (as compare to percutaneous method) to catheter removal was `r round(removalsurvweimod_method_res$HR, 2)[,1]` (95% CI = `r round(removalsurvweimod_method_res$HR, 2)[,2:3]`) times.

```{r}
WeibullDiag(Surv(RemovalDur, CatheterStatus == "Catheter Removed") ~ Method, data = pdcds1)
```

Weibull model seems unfit for method variable, as the lines were linear but not parallel.

```{r}
removalsurv_method_kmfit <- tidy(survfit(Surv(RemovalDur, CatheterStatus == "Catheter Removed") ~ Method, data = pdcds1))
ggplot(removalsurv_method_kmfit, aes(x= log(time), y = log(estimate/(1-estimate)), colour = strata)) +
  geom_point() + geom_line() +
  theme_bw() + ggtitle("PH Assumption Checking Plot - Catheter Removal") + theme(legend.position = c(0.8, 0.8))
```

The plot seems straight line and parallel, thus does not violated proportional hazard assumptions.


### Multivariable

Parametric Survival Analysis with Weibull model was done to find the difference between method, when adjusted to gender, age and diabetes mellitus status.

```{r}
removalsurvweimod <- survreg(Surv(RemovalDur, Compx == "Yes") ~ Gender + Age + CMDM + Method, data = pdcds1, dist = "weibull")
summary(removalsurvweimod)
ConvertWeibull(removalsurvweimod, conf.level = 0.95)
```

when adjusted to gender, age and diabetes status, the method of catheterization had significant effect on catheter survival, with time ratio (TR) of peritonesocopy method to catheter removal was `r round(removalsurvweimod_res$ETR, 2)[4,1]` (95% CI = `r round(removalsurvweimod_res$ETR, 2)[4,2:3]`) times longer than the times need for catheter removal for percutaneous method. 

the hazard ratio (HR) of peritonesocopy method (as compare to percutaneous method) to catheter removal was `r round(removalsurvweimod_res$HR, 2)[4,1]` (95% CI = `r round(removalsurvweimod_res$HR, 2)[4,2:3]`) times , when adjusted to gender, age and diabetes status.

### Prediction

Prediction for Survival Time - Catheter Removal

```{r}

removal_predmod <- predict(removalsurvweimod, newdata = newds, type = "quantile", p = c(.25,.5,.75))
data.frame(cbind(newds, removal_predmod)) %>%
  mutate("Predicted Removal Time (Days) - Median" = X2,
         "Predicted Removal Time (Days) - Q1" = X1,
         "Predicted Removal Time (Days) - Q3" = X3) %>%
  select(Gender, Age, CMDM, Method, "Predicted Removal Time (Days) - Q1", "Predicted Removal Time (Days) - Median", "Predicted Removal Time (Days) - Q3")
```

```{r}
#augment(cxsurvweimod, newdata = pdcds1_ted, type.predict="response") %>% select(Gender, CMDM, Method, CxDur, .fitted)
```

Prediction for Survival Probability

# Discussion

- non-parametric survival - show that for catheter removal, there is significant with percutaneous have higher risk 
- discuss weibull vs other distribution

- gender not significant

- dm significant - impaired wound healing, higher risk to infection

# Conclusion

# Reference


# Additional Info

```{r}

```




```{r}
sessionInfo()
```

